Background for transient faults in asynchronous circuits

### 2019-04-05 draft_together3.tex
editing proof with better english part 1

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\section{Normal operation analysis}
As mentioned in the previous section on Speed Independent circuits, circuits can be modeled by a state graph.  
As the circuit follows the state graph, there may be multiple transitions per state and each path through the states and transitions may be different.  We want to record the succession of states and transitions and so we define a trace as follows:
\begin{definition}A {\em trace} $\sigma$ of a circuit is a sequence of states of infinite length where $\sigma_i$ is the $i^{th}$ term of the sequence.  Formally this is written as $\sigma: \mathbb{N} \to S$ and $T(\sigma_i,\sigma_{i+1})$ for $\forall i \in \mathbb{N}$.\end{definition} 
%add finite traces
% all states or just the state graph ones

We want to define a notion of equivalence between two cicrcuits.  What we observe are the succession of signals which form the trace.  Thus we define circuits as equivalent if the traces are the same. 
\begin{definition} Two circuits A and B are {\em equivalent} if the set of all traces from circuit A is equal to circuit B.  In other words: $\{\sigma^A |\sigma^A$ is a possible trace in $A\}=\{\sigma^B |\sigma^B$ is a possible trace in $B\}$.
\end{definition}
%good enough for now, do I need if and only if?

@@

We make a fault tolerant circuit by the following circuit transformation: make a duplicate copy of the logic blocks to each wire.  We label the duplicated circuit halves as circuit A and circuit B.  Then for each output wire $w$ we connect the output of the logic blocks from each circuit half to three layers of c-elements as shown in figure \ref{fig:dupscheme}.  We label the intermediate wires as $c1^A$, $c2^A$, $c3^A$ and $c1^B$, $c2^B$, $c3^B$ for circuits A and B respectively.  In the first layer of C-elements, $c1^A$ and $c1^B$ are outputs of the logic blocks and are inputs to two C-elements producing $c2^A$ and $c2^B$.  These are then the inputs to the second layer of C-elements producing $c3^A$ and $c3^B$.  These are then the inputs to the third layer of C-elements producing $w^A$ and $w^B$.  We can (arbitrarily) label one of the final C-element outputs as $w^A$ and the other output as $w^B$.  These then connect to other gates in circuit A and circuit B respectively.  \\%(Note that the outputs of the C-elements can be arbitrarily named and connected to the next gates, and the next gates can also be arbitrarily named $A$ or $B$ as long as it does not create any inconsistencies in the naming convention, ie. all $A$ wires feed into the $A$ gates)  %the only wires that really matter are the wires that have isochronic forks, these must input to same A or B, everything else is just a naming difference
\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{circuitforproof3}
  \caption{Duplicating the original logic for each output wire and adding 3 layers of C-elements}
  \label{fig:dupscheme}
\end{figure}

Because the transformed circuit has some extra wires, we want to define a correspondence to compare wires of the original circuit with those of the fault tolerant circuit halves.  A state of the fault tolerant circuit has bits for 8 times as many wires as a state in the original circuit.  Half of those bits are obviously from the A half of the circuit and half are from the B half.  The state graph for a fault tolerant circuit is defined on this combined state.  We define the half-circuit mapping $h^A: s^{\mathit{full}} \to s^A$ where $s^{\mathit{full}}$ is the combined state of a fault tolerant circuit and $s^A$ are made up of the wires $w^A$ that correspond to the original wires $w$.  The half-circuit mapping $h^B$ is similarly defined to map to $s^B$.  We extend $h^A$ and $h^B$ pointwise to traces, in other words, $h$ is applied individually to each state in $\sigma$.  %not sure if I want the mapping to be from variable to variable or specific state to state
For a trace in the fault tolerant circuit we project the full states onto a shortened list of wires (ie the wires corresponding to the original circuit). The resulting trace may have successive repeated states and we delete the repeats to obtain a valid trace on the shortened wires.  %do I need to define all the new transitions in duplicated circuit?  maybe
\\

We want to show this circuit behaves similar to the original circuit under normal operation.  In other words we want to prove that after applying either mapping of $h^A$ or $h^B$, the resultant circuit and the original circuit are equivalent.  We first prove lemma 1 which states for each half of a circuit, if a transition occurs it must follow the transitions of the state graphs in the original circuit.  However this does not guarantee that such a transition always exists so we also need to prove lemma 2 to show that a transition always exists and the circuit cannot deadlock.  These two properties combined together show that all traces in the duplicated circuit can be mapped to a trace in the original circuit.\\

Note that there are some extra wires in the logic blocks in both the original and duplicated circuit.  We actually don't worry about the exact values of these wires other than that if the inputs follow the semi-modularity property then the logic block as an entirety behaves with the semi-modularity property(as long as it is initialized appropriately).  Here it does not suffice for the inputs to be a sequence of valid states and state transitions in the state graph.  This is because the semi-modularity property of the logic block is that if the logic block output is excited then it must remain excited until a transition occurs.  Since certain inputs, even if they are valid states, may cause the logic block to become unexcited before a transition occurs, we want a more stringent criteria.  The inputs must uphold the semi-modularity property which waits at certain input states until the output transitions, in addition to being valid states and state transitions from the state graph.  Then if the wires within the logic block is consistent with Semi-Modularity, that is if the inputs follow Semi-Modularity then there are no glitches at the output and if the output is excited it remains excited until the output changes. \\%newly added

The initial state in the original circuit must be assigned wire values that is consistent with semi-modularity.  For all initial states we assume that we know the values of every wire in the circuit, intermediate and visible wires.  The visible wire values are the output wires in the circuit $w$, and the wires $c1$, $c2$ and $c3$'s. The intermediate wires are all other wires, which consists of all the wires within the logic blocks for each output wire.  Given these we will show how to define a corresponding initial state in the duplicated circuit.  %should I delete intermediate?
\begin{definition} A {\em valid initial state} for the A half of a duplicated circuit is defined as follows.  Given any initial state in the original circuit $s \in S$, assume that for each wire $w$, the wire values within the logic blocks are consistent with the Semi-Modularity property.  Assign the wire values within the logic block to corresponding wires in the half duplicated circuit.  Then assign the value $w$ to corresponding wires in the half duplicated circuit so that $w=c1^A=c2^A=c3^A=w^A$. The same assignment of wire values can also be applied for the B half of a duplicated circuit.
\end{definition}
%\textbf{
For all subsequent discussion in this document, when we talk about the trace of a duplicated circuit we assume that both sides start from the same valid initial state.  I will not include this assumption in the following lemma definitions to avoid complicated wording.  %}  
However for Lemma 1 we only require one half to start from a valid initial state.  
One additional assumption we make about the gates and wire values (including in the presence of faults) are that wire values are strictly 0 or 1, and transitions from excited gates also only produce 0 or 1 and no in between values, no matter the duration of signals from the incoming wires.\\

>>>>

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\section{Normal operation analysis}
As mentioned in the previous section on Speed Independent circuits, circuits can be modeled by a state graph.  
We want to define a trace, to describe the different paths a circuit can take as it follows the state graph.  Different paths can occur because there are states with multiple transitions from that state.  We record the succession of states and transitions by the trace which is defined as follows:
\begin{definition}A {\em trace} $\sigma$ of a circuit is a sequence of states of infinite length where $\sigma_i$ is the $i^{th}$ term of the sequence.  Formally this is written as $\sigma: \mathbb{N} \to S$ and $T(\sigma_i,\sigma_{i+1})$ for $\forall i \in \mathbb{N}$.\end{definition} 
Even if there is only one transition from each state of a state graph, that is a linear state graph, our definition of trace is consistent with that.  \\
%add finite traces
% all states or just the state graph ones

[If the set of traces are equal then the state graph is equal]

Given two circuits, how can we determine when their basic functionality is the same even if they are implemented with different gates and circuit elements?  We define a notion of equivalence between two circuits.  We want the underlying state graph representation of the circuits to be equivalent.  However only the signals and the sequence of its transitions from the circuit is observable.  In other words the trace is observable.  Thus we define circuits as equivalent if the set of traces of the two circuits are the same. 
\begin{definition} Two circuits A and B are {\em equivalent} if the set of all traces from circuit A is equal to circuit B (up to a permutation of the variable names).  In other words, if it is possible to pair wires in A and B in such a way that $\{\sigma^A |\sigma^A$ is a possible trace in $A\}=\{\sigma^B |\sigma^B$ is a possible trace in $B\}$.
\end{definition}
%good enough for now, do I need if and only if? <-NO, xxx if xxx in math lingo is iff

@@

We make a circuit fault tolerant by the following transformation on an existing semi-modular circuit: make a duplicate copy of the logic blocks of each wire $w$.  We label the duplicated logic blocks as belonging to circuit A or circuit B.  Then we connect the output of the logic blocks to three layers of C-elements as shown in figure \ref{fig:dupscheme}.  We label the intermediate wires as $c1^A$, $c2^A$, $c3^A$ and $c1^B$, $c2^B$, $c3^B$ for circuits A and B respectively.  In the first layer of C-elements, $c1^A$ and $c1^B$ are outputs of the logic blocks and are inputs to two C-elements outputting $c2^A$ and $c2^B$.  These are then the inputs to the second layer of C-elements outputting $c3^A$ and $c3^B$.  These are then the inputs to the third layer of C-elements outputting $w^A$ and $w^B$.  We can label one of the final C-element outputs as $w^A$ and the other output as $w^B$.  These then connect to other logic blocks in circuit A and circuit B respectively.  Note that at some steps of the labelling which side is A and B can be somewhat arbitrary because the outputs of pairs of C-elements with the same inputs are symmetric.  The symmetry is broken by connecting the final output $w^A$ to all its circuit A gate inputs and $w^B$ to all its circuit B gate inputs.  All circuit elements with the A label is part of the A-half circuit and all circuit elements with the B label is part of the B-half circuit.\\%(Note that the outputs of the C-elements can be arbitrarily named and connected to the next gates, and the next gates can also be arbitrarily named $A$ or $B$ as long as it does not create any inconsistencies in the naming convention, ie. all $A$ wires feed into the $A$ gates)  %the only wires that really matter are the wires that have isochronic forks, these must input to same A or B, everything else is just a naming difference
\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{circuitforproof3}
  \caption{Duplicating the original logic for each output wire and adding 3 layers of C-elements}
  \label{fig:dupscheme}
\end{figure}

Because the circuit transformation introduces extra wires, we want to define a correspondence to compare states of the original circuit with those of the A-half and B-half of the fault tolerant circuit.  A state of the fault tolerant circuit has bits for 8 times as many wires as a state in the original circuit.  Half of those bits are obviously from the A half of the circuit and half are from the B half.  The state graph for a fault tolerant circuit is defined on this combined state.  We define the half-circuit mapping $h^A: s^{\mathit{full}} \to s^A$ where $s^{\mathit{full}}$ is the combined state of a fault tolerant circuit and $s^A$ consists of wires $w^A$ that correspond to the original wires $w$ in the original duplication procedure.  The half-circuit mapping $h^B$ is similarly defined to map $s^{\mathit{full}}$ to $s^B$.  The added wires such as $c1^A$, $c1^B$, $c2^A$, $c2^B$, $c3^A$, $c3^B$ from the duplication procedure are ignored.  We extend $h^A$ pointwise to traces, in other words, $h^A$ is applied individually to each state in a trace $\sigma$, and likewise for $h^B$.  %not sure if I want the mapping to be from variable to variable or specific state to state
%For a trace in the fault tolerant circuit we project the full states onto a shortened list of wires (ie the wires corresponding to the original circuit). 
The resulting trace may have successive repeated states and we delete the neighboring repeats to obtain a valid trace on the shortened list of wires.  %do I need to define all the new transitions in duplicated circuit?  maybe
\\

To prove claims about the fault tolerance of this circuit, we must first prove that the circuit behaves similar to the original circuit under normal operation.  That is, after applying a mapping of $h^A$, the set of traces of the resultant circuit and that of the original circuit are equal.  And likewise after applying a mapping of $h^B$.  We prove Lemma 1 which shows for each half of a circuit, if a transition occurs it must follow the transitions of the state graphs in the original circuit.  This property is independent of the signals on the other half circuit and is useful later in proving fault tolerance under a single transient fault.  %as long as it does not violate digital abstraction
However Lemma 1 does not guarantee that such a transition always exists so Lemma 2 is needed to show that a transition always exists and as a consequence the circuit cannot deadlock.  These two Lemmas combine together to show that all traces in the duplicated circuit can be mapped to a trace in the original circuit.\\% and all traces in the original can be mapped to duplicated after h mapping?  <-this is obvious though by setting the A half and B half to both transition according to original trace

The internal wires in the logic blocks that are not included in the state graphs need to have values that are semi-modularity consistent for the circuit to proceed with no glitches.  The internal wires exist in the logic blocks in the original circuit and thus exist in the fault tolerant circuit from the circuit transformation.  The actual values of these wires do not matter if the inputs to the logic blocks follow the semi-modularity property, and if the internal wires are initialized properly.  This arises from the original synthesis method to synthesize the circuit from an underlying state graph.  
Additionally, it does not suffice for the inputs to the logic blocks to be from a valid trace of the state graph; the timing matters as well.  The semi-modularity property must be preserved:  if the logic block output is excited then it must remain excited until a transition on its output occurs.  Since certain inputs which are valid states and arise from valid transitions may cause the logic block to become unexcited before a transition occurs and thus break the semi-modularity property, we require this more stringent criteria.  %maybe needs an example
Thus to satisfy semi-modularity, the inputs wait at certain input states until the output transitions.  Then the internal wires of the logic block are termed {\em semi-modularity consistent} if the following property is true:  if the inputs follow semi-modularity then there are no glitches at the output and that once the output is excited it remains excited until it transitions. \\%newly added

As mentioned in the previous paragraph, the internal wires The initial state in the original circuit must be assigned wire values that is consistent with semi-modularity.  For all initial states we assume that we know the values of every wire in the circuit, intermediate and visible wires.  The visible wire values are the output wires in the circuit $w$, and the wires $c1$, $c2$ and $c3$'s. The intermediate wires are all other wires, which consists of all the wires within the logic blocks for each output wire.  Given these we will show how to define a corresponding initial state in the duplicated circuit.  %should I delete intermediate?
\begin{definition} A {\em valid initial state} for the A half of a duplicated circuit is defined as follows.  Given any initial state in the original circuit $s \in S$, assume that for each wire $w$, the wire values within the logic blocks are consistent with the semi-modularity property.  Assign the wire values within the logic block to corresponding wires in the half duplicated circuit.  Then assign the value $w$ to corresponding wires in the half duplicated circuit so that $w=c1^A=c2^A=c3^A=w^A$. The same assignment of wire values can also be applied for the B half of a duplicated circuit.
\end{definition}
%\textbf{
For all subsequent discussion in this document, when we talk about the trace of a duplicated circuit we assume that both sides start from the same valid initial state.  I will not include this assumption in the following lemma definitions to avoid complicated wording.  %}  
However for Lemma 1 we only require one half to start from a valid initial state.  
One additional assumption we make about the gates and wire values (including in the presence of faults) are that wire values are strictly 0 or 1, and transitions from excited gates also only produce 0 or 1 and no in between values, no matter the duration of signals from the incoming wires.\\
